Congestion-Aware Join Optimization in Key-Value Stores

### 2014-08-07 paper/paper.tex
Report edits

For example, Xiong et al.~\cite{Xiong2014} examine the usefulness of software-defined networking in supporting distributed analytical queries.
Their workload consits of read-only SQL queries where query processing is bandwidth-intensive.
They construct a global query optimizer which decides on join order and how query results should be passed between sites.
This work takes a similar approach to ours, but also identifies many opportunities for further work.

@@

>>>>

For example, Xiong et al.~\cite{Xiong2014} examine the usefulness of software-defined networking in supporting distributed analytical queries.
Their workload consists of read-only SQL queries where query processing is bandwidth-intensive.
They construct a global query optimizer which decides on join order and how query results should be passed between sites.
This work takes a similar approach to ours, but also identifies many opportunities for further work.

@@

Each experiment is performed 5 times with the results are used to compute the data in the following sections.
The 95th percentile confidence interval is shown where appropriate.

\subsection{Micro Benchmark}
\begin{figure}
    \centering
    \includegraphics[scale=2.5]{figures/Join_micro_benchmark.png}
    \caption{Micro benchmark setup}\label{fig:micro_benchmark}
\end{figure}

First, we want to demonstrate the proof-of-concept of our design by evaluating our system with the simple topology shown in Figure~\ref{fig:micro_benchmark}.
We engineered the client query such that each server holds a different portion of the relevant data for the join operation.
That is, server 1 holds $1/4$ of the relevant item data and server 2, 3, and 4 hold $1/16$, $1/32$, $1/64$ of the relevant item data respectively.
The total sizes of the item and user data are approximately 9 MB and 2 MB\@.
In order to satisfy our assumption where the network is the bottleneck, we set the capacity of each link in the topology to 1 Mbps.
The client is connected directly to the core switch with a link of the same capacity.

In Figure~\ref{fig:micro_benchmark_total_time}, the red bars represent our system's baseline performance without our scheduling mechanism is turned off.
The average completion time without scheduling for each client query is over 180 seconds and also shows significant variation.
To provide predictable performance with our scheduling mechanism, we must first understand the characteristics of the costs for each component of the join operation.
In our case it is easy to predict the costs because we assigned different portions of relevant data to each server.
However, it is also easy in a real deployment scenario as the data distribution among servers is typically available or can be easily computed.
Recall that by design, server 1 is responsible for transmitting the largest flows between other servers and the client. 
Therefore, our scheduling must take this information into consideration and ensure that server 1 has the largest minimum bandwidth guarantee for this particular operation.

Though we want server 1 to transfer its data as fast as possible, we do not want to starve other servers.
Therefore, we must also have reasonable minimum bandwidth guarantees for other servers.
Through experimentation, we found the most suitable minimum bandwidth guarantees on the 1 Mbps link for servers 2, 3, and 4 are 0.3 Mbps, 0.2 Mbps and 0.1 Mbps respectively.
We fixed these parameters and varied the minimum bandwidth guarantee for server 1 to study the network behaviour.
As shown by the right bars in Figure~\ref{fig:micro_benchmark_total_time}, there is a noticeable performance gain when we increase the minimum bandwidth guarantee of server 1 from 0.5 Mbps to 0.8 Mbps.
This shows that it is important to favour larger flows during a join operation to minimize the overall completion time.
We also want to highlight that we see diminishing returns as we further increase the guarantee.
This is caused by a significant increase in the completion time for other flows, hurting the overall performance.
In the best case, with 0.8 Mbps minimum bandwidth guarantee for server 1, the average completion time of the client query is 154 seconds which gives us almost 15\% performance gain over the baseline implementation.

\begin{figure}
    \centering
    \includegraphics[scale=0.47]{figures/simple_topo_overall.png}
    \caption{Query completion time}\label{fig:micro_benchmark_total_time}
\end{figure}

Next, we further examine the performance results to understand why network scheduling helps.
Figure~\ref{fig:sub_query_completion_8} shows the subquery completion time for each server.
The per-server completion is calculated by the client after it fetches all the required data from each server.
The bars on the left are generated when traffic scheduling is disabled.
We see that server 1 takes the most time to complete and server 4 takes the least time.
This is expected due to the data distribution we defined above.
We can also see that when we enabled our network traffic scheduling mechanisms, the completion time for server 3 and server 4 slowed down slightly.
This is because most of the bandwidth will be allocated to help server 1 to finish its task as soon as possible.
However, The end result is that traffic scheduling reduces the average subquery completion time, which in turn reduces the overall query execution time.
We see that our experimental results match with our expectations.

\begin{figure}
    \centering
    \includegraphics[scale=0.47]{figures/simple_topo_closer_8.png}
    \caption{Sub query completion time}\label{fig:sub_query_completion_8}
\end{figure}
